---
title: "script1"
output:
  pdf_document: default
  html_document: default
---

##Read data
```{r}
library(dplyr)
yr2015<-read.csv("2015_world_happiness.csv")
yr2016<-read.csv("2016_world_happiness.csv")
yr2017<-read.csv("2017_world_happiness.csv")

yr2015<-yr2015[,c(1:4,6:12)]
yr2016<-yr2016[,c(1:4,7:13)]
yr2017<-yr2017[,c(1:3,6:12)]

regiondata<-yr2016[,c(1,2)]
yr2017<-left_join(yr2017,regiondata,by = "Country")
yr2017<-yr2017[,c(1,11,2:10)]

yr2015$rank<-1:nrow(yr2015)
yr2016$rank<-1:nrow(yr2016)
yr2017$rank<-1:nrow(yr2017)
```

##Merge and caculate the difference in rankings between 2017 and 2015
```{r}
countries<-left_join(yr2015[,c(1,12)],yr2016[,c(1,12)],by="Country")
countries<-left_join(countries,yr2017[,c(1,12)],by="Country")
colnames(countries)[c(2,3,4)]<-c("rank2015","rank2016","rank2017")

countries$delta<-countries$rank2017-countries$rank2015
countries<-countries[complete.cases(countries),]
summary(countries$delta)

```

**1st Quantiles is -10. 3rd Quantile is 5.**

##For all countries that have moved up more than 10 places in their rankings from 2015 to 2017
```{r}
improve<-countries %>% filter(delta < -10)

getcountry <- function(country){
  df<-yr2015 %>% filter(Country == country) %>% cbind(yr2017 %>% filter(Country == country))
  return(df)
}
countryimprove<-improve$Country
improved_set<-getcountry(countryimprove[1])
for (i in 1:length(countryimprove)){
  improved_set<-rbind(improved_set,getcountry(countryimprove[i+1]))
  
}
improved_set$gdpdiff <- improved_set[,17]-improved_set[,5]
improved_set$familydiff<-improved_set[,18]-improved_set[,6]
improved_set$healthdiff<-improved_set[,19]-improved_set[,7]
improved_set$freedomdiff<-improved_set[,20]-improved_set[,8]
improved_set$trustdiff<-improved_set[,22]-improved_set[,9]
improved_set$gendiff<-improved_set[,21]-improved_set[,10]
improved_set$dysdiff<-improved_set[,23]-improved_set[,11]

summary(improved_set[,c(25:31)])
```

**If only look at the mean, some indicators are still negative. Seems like the most influential factors are family, gdp and dystopia residual.**

## 5 Countries improved the most
```{r}
Latvia<-improved_set[improved_set$Country == "Latvia",c(25:31)]
Egypt<-improved_set[improved_set$Country == "Egypt",c(25:31)]
Romania<-improved_set[improved_set$Country == "Romania",c(25:31)]
Hungary<-improved_set[improved_set$Country == "Hungary",c(25:31)]
Bulgaria<-improved_set[improved_set$Country == "Bulgaria",c(25:31)]

fiveimprovecountry<-c("Latvia","Egypt","Romania","Hungary","Bulgaria")
fiveimprove<-rbind(Latvia, Egypt, Romania,Hungary, Bulgaria)
fiveimprove$country<-fiveimprovecountry
fiveimprove<-fiveimprove[,c(8,1:7)]
fiveimprove
```



##For all countries that have dropped more than 5 places in their rankings from 2015 to 2017

```{r}
drop<-countries %>% filter(delta >5)
countrydrop<-drop$Country
drop_set<-getcountry(countrydrop[1])
for (i in 1:length(countrydrop)){
  drop_set<-rbind(drop_set,getcountry(countrydrop[i+1]))
  
}

drop_set$gdpdiff<-drop_set[,17]-drop_set[,5]
drop_set$familydiff<-drop_set[,18]-drop_set[,6]
drop_set$healthdiff<-drop_set[,19]-drop_set[,7]
drop_set$freedomdiff<-drop_set[,20]-drop_set[,8]
drop_set$trustdiff<-drop_set[,22]-drop_set[,9]
drop_set$gendiff<-drop_set[,21]-drop_set[,10]
drop_set$dysdiff<-drop_set[,23]-drop_set[,11]
summary(drop_set[,c(25:31)])


```


##Country dropped ranking the most
```{r}
Venezuela<-drop_set[drop_set$Country == "Venezuela",c(25:31)]
Liberia<-drop_set[drop_set$Country == "Liberia",c(25:31)]
Zambia<-drop_set[drop_set$Country == "Zambia",c(25:31)]
Haiti<-drop_set[drop_set$Country == "Haiti",c(25:31)]
Zimbabwe<-drop_set[drop_set$Country == "Zimbabwe",c(25:31)]


fivedropcountry<-c("Venezuela","Liberia","Zambia","Haiti","Zimbabwe")
fivedrop<-rbind(Venezuela, Liberia, Zambia,Haiti, Zimbabwe)
fivedrop$country<-fivedropcountry
fivedrop<-fivedrop[,c(8,1:7)]
fivedrop
```


